---
title: "Stratified Sampling"
author: "A.Mayuri(2348133)"
date: "2023-09-12"
output: word_document
---
## Question
Consider a dataset as
population of your choice and divide the population in various strata by
choosing an appropriate variable of stratification.  Give the estimates of population
parameters(mean and total) by taking a random sample of adequate size using the
proportional allocation and optimum allocation methods. Write a report on it.

## Data Description:
The taken dataset gives a detailed description of loan defaults by customers on different kinds of auto-mobiles.
the data is of large sample with N=200 customers in the past week


## Objective: 
To do a stratified analysis on the employee type ie(salaried/self employed)

## Variables of Interest and their Definition:
UniqueID-Identifier for customers
loan_default-- Payment default in the first EMI on due date
disbursed_amount--Amount of Loan disbursed
asset_cost--Cost of the Asset
ltv--Loan to Value of the asset
branch_id--Branch where the loan was disbursed
supplier_id--Vehicle Dealer where the loan was disbursed
manufacturer_id--Vehicle manufacturer(Hero, Honda, TVS etc.)
Employment.Type--Employment Type of the customer (Salaried/Self Employed)

where we will do the analysis for 2 stratum ie, salaried=1, and stratum=0

# ANALYSIS

## Code

### Step1: IMPORT DATASET
To import the dataset that has been discussed above to initiat the analysis
```{r }
library(readxl)
finb <- read_excel("C:/Users/mayur/Desktop/Mstat/tri sem 1/R/dataset/finb.xlsx")
View(finb)
attach(finb)


```
### Step2: IMPORT PACKAGE
We are downloading the samplingbook package to proceed with the stratified sampling data analysis
```{r }
library(samplingbook)
```
### step3: CREATE STRATUMS
Salaried =1 
selfemployed =0

this helps us to create stratums by assigning a binary code for saalaried and self employed

```{r }
# creating stratums . employment type= salaried
stratum1=finb[finb$Employment.Type==1
, ]
stratum1

```



```{r }

# creating stratums . employment type= self employed
stratum2=finb[finb$Employment.Type==0
, ]
stratum2
```

### step4: CALCULATE
mean and standard deviation for two variables disumbered amount and asset cost

```{r }
# calculation for N-stratum population size
N1=sum(stratum1$Employment.Type==1)
N1

N2=sum(stratum2$Employment.Type==0)
N2

# mean calculation
M1_disamount=mean(stratum1$disbursed_amount)
M2_disamount=mean(stratum2$disbursed_amount)

M1_ascost=mean(stratum1$asset_cost)
M2_ascost=mean(stratum2$asset_cost)

#std deviation calculation

S1_disamount=sqrt(var(stratum1$disbursed_amount))
S1_ascost=sqrt(var(stratum1$asset_cost))

S2_disamount=sqrt(var(stratum2$disbursed_amount))
S2_ascost=sqrt(var(stratum2$asset_cost))

#output of mean and standard deviation for two variables disumbered amount and asset cost
M1_disamount
M1_ascost
S1_ascost
S1_disamount



M2_disamount
M2_ascost
S2_ascost
S2_disamount

```

## step5: PROPORTIONAL ALLOCATION

let the sample of size n=10 has to be drawn using proportional allocation


```{r }
#let the sample of size n=10 has to be drawn using proportional allocation for asset cost

sample_size_ascost=stratasamp(n=10, Nh=c(N1, N2), Sh=c(S1_ascost, S2_ascost), type="opt")
sample_size_ascost

#let the sample of size n=10 has to be drawn using proportional allocation for disumbersed amount

sample_size_disamt=stratasamp(n=10, Nh=c(N1, N2), Sh=c(S1_disamount, S2_disamount), type="opt")
sample_size_disamt
```

### Step6: determination of sample size

1)proportional
```{r }
# determination of total sample size for given specified precision using proportion
stratasize(e=.1, Nh=c(N1, N2), Sh=c(S1_ascost, S2_ascost), type='prop' )
stratasize(e=.1, Nh=c(N1, N2), Sh=c(S1_disamount, S2_disamount), type='prop' )
```

2)Optimal

```{r}
# determination of total sample size for given specified precision using proportion
stratasize(e=.1, Nh=c(N1, N2), Sh=c(S1_ascost, S2_ascost), type='opt' )
stratasize(e=.1, Nh=c(N1, N2), Sh=c(S1_disamount, S2_disamount), type='opt' )
```


3) with precision
```{r}
# determination of total sample size for given specified precision and confidence level

stratasize(e=.1, Nh=c(N1, N2), Sh=c(S1_ascost, S2_ascost),level=.99, type="opt" )


stratasize(e=.1, Nh=c(N1, N2), Sh=c(S1_disamount, S2_disamount),level=.99, type="opt" )


```
### Step7: COLLECTION OF RANDOMN DATASET OF Size=5

```{r}
#collect a random sample of size 5,5 from both strata
sample1=stratum1[sample(1:nrow(stratum1), 5, replace=FALSE), ]
sample1 # sample 1 collected from stratum 1

sample2=stratum2[sample(1:nrow(stratum2), 5, replace=FALSE), ]
sample2 # sample 1 collected from stratum 2

# total sample collected using stratified random sampling
total_sampled_data=rbind(sample1, sample2)
total_sampled_data
```


### step8: ESTIMATIONS

```{r}
#with optimum allocation

# Estimation of population mean using stratified random sample

nh2=as.vector(table(total_sampled_data$Employment.Type))
nh2

wh=nh2/sum(nh2)
wh
```

1) Disbursed amount

```{r}
stratamean(y=total_sampled_data$disbursed_amount, h=as.vector(total_sampled_data$Employment.Type),
wh=wh, eae=TRUE)


stratamean(y=total_sampled_data$disbursed_amount, h=as.vector(total_sampled_data$Employment.Type),
wh=wh)
```
Interpretation: 

#### General interpretation:: 

Thus by using stratified sampling we have deduced the  mean estimate of disbursed amount is of 49,212 rupees with a standard deviation of 1,590 rupees. also the mean estimate lied between [46095.51,52329.69] with a 95% confidence level.

Salaried professionals have a mean estimate of   46,519.6 rupees as their disbursed amount with a std error of 2277.471 rupees.

Self employed have a mean estimate of  51,905.6 rupees as their disbursed amount with a std error of 2220.445 rupees.


# Conclusion :: 
Here we can observe that self employed people will default more than that of salaried. this could be due to the uncertainty in income, which is higher in self employed individuals.


2) Asset cost

```{r}
stratamean(y=total_sampled_data$asset_cost, h=as.vector(total_sampled_data$Employment.Type),
wh=wh, eae=TRUE)


stratamean(y=total_sampled_data$asset_cost, h=as.vector(total_sampled_data$Employment.Type),
wh=wh)
```
Interpretation: 

####General interpretation:: 

Thus by using stratified sampling we have deduced the  mean estimate of asset cost is of 66,294.4 rupees with a standard deviation of 1,426.502 rupees. also the mean estimate lied between [63,498.51,69,090.29] with a 95% confidence level.

Salaried professionals have a mean estimate of  68,050.2  rupees as their asset amount with a std error of 2,647.529 rupees.

Self employed have a mean estimate of 64,538.6 rupees as their asset amount with a std error of 1,063.117rupees.


### Conclusion :: 
Here we can observe that self employed people will take loan of lower value than that of salaried. this could be due to the uncertainty in income, which is higher in self employed individuals.

